---
title: "Identifying and Understanding key factors of sexual and reproductive health for Chinese female college students"
author: "Zongchao Liu; Zhi Lin"
date: "1/10/2021"
output: 
  html_document:
          toc: true 
          toc_float: true
          code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arsenal)
library(knitr)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(caret)
library(ranger)
library(lime)
options(java.parameters = "-Xmx2500m")
library(bartMachine)
library(gplots)
library(factoextra)
library(pROC)
library(ResourceSelection)
library(conflicted)
conflict_prefer("select", "dplyr", "MASS")
conflict_prefer("filter", "dplyr", "stats")
conflict_prefer("rename", "dplyr")
```

# Intro

This report is the second project for a SRH project. The main goal of this project is to identify and understand key factors of adverse outcomes for Chinese female college students who are considered to be sexually active.

# Import & cleaning data

```{r message=FALSE, warning=FALSE}
NCSS_SRH = read_csv('./data/2020NCSS-SRH_0228cleaned.csv')
var_names = names(NCSS_SRH) %>% as.tibble()

attitude = NCSS_SRH %>%
  filter(!is.na(`B24.避孕措施使用频率code`)) %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  select(15, seq(133, 155, 2)) 
  #filter(`P6.生理性别code` == 0)

att_list = names(attitude)
names(attitude[,-1]) = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12")

awareness = NCSS_SRH %>%
  select(14,70,242,111,113:121)

harass = NCSS_SRH %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  #filter(`P6.生理性别code` == 0) %>%
  select(1,389:395) %>%
  mutate(`B43.(性骚扰)因身体发育被同学嘲笑(可多选)` = ifelse(str_detect( `B43.(性骚扰)因身体发育被同学嘲笑(可多选)`, "上大学"), 1, 0),
         `B43.(性骚扰)在现实生活中受到言语上的性骚扰(可多选)` = ifelse(str_detect( `B43.(性骚扰)在现实生活中受到言语上的性骚扰(可多选)`,"上大学"), 1, 0),
         `B43.(性骚扰)在网上或手机上受到言语上的性骚扰(可多选)` = ifelse(str_detect(`B43.(性骚扰)在网上或手机上受到言语上的性骚扰(可多选)`,"上大学"), 1, 0),
         `B43.(性侵害)强迫脱衣/暴露隐私部位(可多选)` = ifelse(str_detect(`B43.(性侵害)强迫脱衣/暴露隐私部位(可多选)`,"上大学"), 1, 0),
         `B43.(性侵害)强迫亲吻或触摸隐私部位(可多选)` = ifelse(str_detect(`B43.(性侵害)强迫亲吻或触摸隐私部位(可多选)`,"上大学"), 1, 0),
         `B43.(性侵害)强迫口交(可多选)` = ifelse(str_detect( `B43.(性侵害)强迫口交(可多选)`,"上大学"), 1, 0),
         `B43.(性侵害)强迫性交(阴道/肛门)(可多选)` = ifelse(str_detect( `B43.(性侵害)强迫性交(阴道/肛门)(可多选)`,"上大学"), 1, 0))

colSums(harass[,-1])
sum(rowSums(harass[,-1]) != 0)

outcomes = NCSS_SRH %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  #filter(`P6.生理性别code` == 0) %>%
  select(1, 274, 388, 511) %>%
  rename("contraceptive_freq" = "B24.避孕措施使用频率code",
         "std" = "B42.是否确诊过性病",
         "depression" = "C10.抑郁量表CES-D-10总分(≥10可认为有抑郁症状)") %>%
  mutate(contraceptive_freq = as.numeric(contraceptive_freq),
         contraceptive_freq = case_when(contraceptive_freq == 5 ~ 3,
                                         contraceptive_freq == 4 ~ 2,
                                        contraceptive_freq == 3 | contraceptive_freq == 2 | contraceptive_freq == 1 ~ 1),
         depression = factor(ifelse(depression >= 10, 1, 0)),
         harass = factor(ifelse(rowSums(harass[,-1]) != 0, 1, 0))) %>%
  #mutate(pregnant = ifelse(is.na(pregnant), 0, pregnant)) %>%
  rename("key" = "序号")

# colSums(outcomes)
p = NCSS_SRH[,273:274] %>% na.omit()
p %>% filter(`B24.避孕措施使用频率code` == 2)
sum(p[,2] == 5)
sum(p[,2] == 4)
sum(p[,2] == 3)
sum(p[,2] == 2)
sum(p[,2] == 1)
```

## Outcomes

In total, we have 6968 recored subjects who have hookup experience.

```{r}
head(outcomes) %>% kable()
```


The adverse outcomes we define here include:

- unwanted pregnancy: Among the 2199 subjects, we have `r sum(outcomes$pregnant == 1)` subjects who have the outcome of unwanted pregnancy

- STD: Among the 2199 subjects, we have `r sum(outcomes$STD == 1)` subjects who have the outcome of STD

- Sexual assault: Among the 2199 subjects, we have `r sum(rowSums(harass[,-1]) != 0)` subjects who have the outcome of Sexual assault

- Depression: Among the 2199 subjects, we have `r sum(outcomes$depression.y == 1)` subjects who have the outcome of depression

## Covariates

- School related:

location, school type, undergraduate, enrollment year, education level(index: `7,8,9,11,12,13`)

```{r}
school = NCSS_SRH %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  select(15,1,7,8,9,11,12,13) %>%
  #filter(`P6.生理性别code` == 0) %>%
  select(-1)
```


- SES

sex, age, expectation for marriage, expenditure, romantic relationship, father/mother's education level*, smoking, alcohol consumption, single child, area, divorced family

```{r}
demographics = NCSS_SRH %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  #filter(`P6.生理性别code` == 0) %>%
  select(c(1,14,19,38,519,159,534,536,469,467,131,242,533,511,525, 556)) %>% # 242 indicates sexual com
  rename(expenditure =`D4.过去12个月平均每月支出(元,不含学校规定缴费项目)`,
         age = `实足年龄(周岁,取整)`,
         divorce = `D16.家中是否发生过父母离异`) %>%
  mutate(expenditure = case_when(expenditure <= 1500 ~ "<1500",
                                 expenditure <= 2500 & expenditure > 1500 ~ "1500~2500",
                                 expenditure > 2500 ~ ">2500"),
         #age = case_when(age < 20 ~"<20",
         #                age >=20 & age<= 21 ~ "21~22",
         #                age > 22 ~ ">22"),
         divorce = factor(divorce))
```

- Attitudes

we synthesized 12 questions

```{r}
attitude = attitude[,-1]
#attitude[attitude == 1 | attitude == 2] = -1
#attitude[attitude == 3 ] = 0
#attitude[attitude == 4 | attitude == 5] = 1
hclust.complete = hclust(dist(attitude[,-1]), method = "complete")
plot(hclust.complete)
ind3.complete = cutree(hclust.complete, 3)
```


- Awareness(SRH knowledge) 

Initially we can use the SRH score for predicting the outcomes

```{r}
awareness = NCSS_SRH %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  #filter(`P6.生理性别code` == 0) %>%
  select(1,131) %>%
  rename("score" = "A16.性知识得分(满分9)")
```

- Contraceptive

The person who determine the contraceptive use; Accessibility; free access

indices are: `326, 328, 330(331)`

```{r}
contraceptive = 
  NCSS_SRH %>%
  filter(`B7.是否发生过插入式性行为(阴道/肛门)` == 1 ) %>%
  #filter(`P6.生理性别code` == 0) %>%
  select(1,325, 327, 331)
  
unique(contraceptive$pdeter)
unique(contraceptive$accessibility)
names(contraceptive) = c("key", "pdeter", "accessibility", "freeacc")

contraceptive = contraceptive %>%
  mutate(accessibility = factor(ifelse(str_detect(accessibility, "不"), 0, 1)),
         freeacc = factor(freeacc)) %>%
  filter(!str_detect(pdeter, "其他"))
```


Combined covariates: 

```{r}
covariates = left_join(left_join(school, demographics),awareness) %>% rename("key" = "序号")
names(covariates) = c("key", "location", "schlevl","schtype","degree","undergrad","enrllyr","sex","age","marriage","expenditure","relationship","fedu","medu","smoking","drink","srh_score", "hookup", "single","depression", "area", "divorce", "score")
```



# General Workflow

- feature engineering

```{r}
features = covariates %>% 
  mutate(location = case_when(location == 1 ~ "east",
                              location == 2 ~ "middle",
                              location == 3 ~ "west"),
         schlevl = case_when(schlevl == "本科" ~ "univ",
                             schlevl == "专科" ~ "vocational"),
         schlevl = factor(schlevl, levels = c("vocational", "univ")),
         degree = factor(degree),
         undergrad = factor(undergrad),
         enrllyr = enrllyr - 2015,
         sex = case_when(sex == "男" ~ "male",
                         sex == "女" ~ "female"),
         marriage = as.character(marriage),
         expenditure = factor(expenditure, levels = c("<1500", "1500~2500", ">2500")),
         expenditure = case_when(expenditure == "<1500" ~ 0,
                                 expenditure == "1500~2500" ~ 1,
                                 expenditure == ">2500" ~ 2),
         drink = factor(drink),
         single = factor(single),
         fedu = case_when(str_detect(fedu, "大学") ~ "bachelor",
                          str_detect(fedu, "硕士") ~ "master+",
                          !str_detect(fedu, "大学") & !str_detect(fedu, "硕士") ~ "lower"),
         medu = case_when(str_detect(medu, "大学") ~ "bachelor",
                          str_detect(medu, "硕士") ~ "master+",
                          !str_detect(medu, "大学") & !str_detect(medu, "硕士") ~ "lower"),
         fedu = factor(fedu, levels = c("lower", "bachelor", "master+")),
         medu = factor(fedu, levels = c("lower", "bachelor", "master+")),
         area = case_when(area == "农村" ~ "rural",
                          area == "城市" ~ "urban",
                          area == "县城(乡镇、郊区)" ~ "suburban"),
         relationship = case_when(relationship == "目前单身(恋爱过)" ~ "2",
                                  relationship == "有男/女朋友(及已婚)" ~ "3",
                                  relationship == "一直单身" ~ "1"
                                  ),
         smoking = case_when(smoking == "从不吸烟" ~ "1",
                             smoking == "过去吸烟，但目前已不抽" ~ "2",
                             smoking == "现在吸烟" ~ "3")
         
         ) %>%
  select(-srh_score, -hookup, -schtype)
```

- Combine targets with features

```{r}
df = features %>%
  cbind(., ind3.complete) %>%
  rename(attitude = ind3.complete) %>%
  left_join(., outcomes, by = "key") %>%
  left_join(., contraceptive, by = "key") %>%
  select(-1) %>%
  select(everything(), contraceptive_freq, std, depression.y, harass)
```


- Generate specific datasets for each outcome

```{r}
contraceptive_pred = df %>% na.omit() %>%
  mutate(pdeter = case_when(pdeter == "自己" ~ "1",
                            pdeter == "对方" ~ "4",
                            pdeter == "双方协商" ~ "2",
                            pdeter == "说不好谁来决定，看情况" | pdeter == "看手头有什么就用什么" ~ "3")) %>%
  filter(age > 11 & age < 30)
  
#std_pred = df[,-c(21,23,24)] %>% na.omit()
#depression_pred = df[,-c(21,22,24)] %>% na.omit()
#harass_pred = df[,-c(21:23)] %>% na.omit()

```


# Unwanted pregnancy

## split original data

```{r}
#down sampling
t = contraceptive_pred %>% mutate(contraceptive_freq = factor(contraceptive_freq))
down_df = downSample(t, t$contraceptive_freq) %>% mutate(contraceptive_freq = as.numeric(contraceptive_freq)) %>% select(-Class)
contraceptive_pred = down_df %>% mutate(attitude = as.character(attitude),
                                        contraceptive_freq = case_when(contraceptive_freq == 1 ~ "low",
            contraceptive_freq == 2 ~ "middle",
            contraceptive_freq == 3 ~"high"),
            contraceptive_freq = factor(contraceptive_freq, levels = c("low","middle","high"))) %>%
  mutate(contraceptive_freq = as.numeric(contraceptive_freq)) 
# %>% mutate(contraceptive_freq = scale(contraceptive_freq, scale = T))
  

set.seed(123)
trRow = createDataPartition(y = contraceptive_pred$contraceptive_freq, p = 0.7)$Resample1
```


## pregnacy - training

```{r}
x = model.matrix(contraceptive_freq ~ . , contraceptive_pred)[,-1] %>% as.tibble() %>%
  as.matrix() %>% .[trRow, ]
x.lasso.test = model.matrix(contraceptive_freq ~ . , contraceptive_pred)[,-1] %>% as.tibble() %>%
  as.matrix() %>% .[-trRow, ]
y = contraceptive_pred$contraceptive_freq[trRow]

#check dim
dim(x)
length(y)

ctrl = trainControl(method = "cv", number = 5,
                    classProbs = F,
                    #summaryFunction = twoClassSummary,
                    #sampling = "up"
                    )

##lr
#lr.fit = train(x = contraceptive_pred[trRow,-21], # rm std
#                y = contraceptive_pred$contraceptive_freq[trRow],
#                method = "glm",
#                #metric = "ROC",
#                trControl = ctrl
#                ) 
#
##lasso
#lasso.fit = train(x,y,
#                method = "glmnet",
#                #metric = "ROC",
#                tuneGrid = expand.grid(alpha = 1,lambda = exp(seq(-7, -6, length #= 100))),
#                trControl = ctrl)

#ggplot(lasso.fit, highlight = T) +
#  labs(title = "LASSO tuning procedure") +
#  theme_bw()

# rf
set.seed(123)
rf.grid = expand.grid(mtry = 12:17,
                      splitrule = "variance",
                      min.node.size = c(100,150,200)
                      )
rf.fit = train(contraceptive_freq ~ ., contraceptive_pred[trRow,], 
                method = "ranger",
                tuneGrid = rf.grid,
                #metric = "ROC",
                trControl = ctrl)
ggplot(rf.fit, highlight = T)  +
  labs(title = "RF tuning procedure") +
  theme_bw()

# Adaboost 
gbmA.grid = expand.grid(n.trees = c(50,75,100),
                        interaction.depth = 1:10,
                        shrinkage = c(0.1),
                        n.minobsinnode = c(100,150,200))


set.seed(123)
gbmA.fit = train(contraceptive_freq ~ ., contraceptive_pred[trRow,],
                 tuneGrid = gbmA.grid,
                 trControl = ctrl,
                 method = "gbm",
                 distribution = "gaussian",
                 #metric = "ROC",
                 verbose = T)
gbmA.fit$bestTune 
ggplot(gbmA.fit, highlight = T) +
    labs(title = "GBM tuning procedure") +
  theme_bw()


set.seed(123)
rf.final = ranger(contraceptive_freq ~ ., 
                  contraceptive_pred[trRow,], 
                  mtry = rf.fit$bestTune$mtry, 
                  splitrule = "variance",
                  min.node.size = rf.fit$bestTune$min.node.size,
                  importance = "impurity")

# var importance
vip::vip(rf.final,num_features = 23) + theme_bw() + ggtitle("Variable Importance - RF")
vip::vip(gbmA.fit$finalModel,num_features = 23) + theme_bw() + ggtitle("Variable Importance - GBM")


#bart
set.seed(123)
x.bart = contraceptive_pred[trRow,-21] %>% as.data.frame()
x.bart.test = contraceptive_pred[-trRow,-21] %>% as.data.frame()

y.bart = as.integer(contraceptive_pred$contraceptive_freq[trRow])
bart.fit.cv = bartMachineCV(X = x.bart,y = y.bart, k_cvs = 5, verbose = T, num_tree_cvs = c(50,100))

#bart.fit = build_bart_machine(X = x.bart, 
#                             y = y.bart,
#                             num_trees = 150, 
#                             num_burn_in = 200,
#                             num_iterations_after_burn_in = 800)


```


```{r}
############### training auc
#lr.fit.pred = predict(lr.fit,newdata = contraceptive_pred[trRow,-c(21)])
#sum(lr.fit.pred == contraceptive_pred$contraceptive_freq[trRow])/length(lr.fit.pred)

#lasso.fit.pred = predict(lasso.fit,newdata = x)
#sum(lasso.fit.pred == #contraceptive_pred$contraceptive_freq[trRow])/length(lasso.fit.pred)

rf.fit.pred = predict(rf.fit,newdata = contraceptive_pred[trRow,-c(21)])
sum(rf.fit.pred == contraceptive_pred$contraceptive_freq[trRow])/length(rf.fit.pred)
rf_res_train = tibble(pred = as.numeric(rf.fit.pred), true = as.numeric(contraceptive_pred$contraceptive_freq[trRow]))
rf_training_err = sqrt(mean((rf_res_train$true - rf_res_train$pred)^2))

gbmA.fit.pred = predict(gbmA.fit,newdata = contraceptive_pred[trRow,-c(21)])
gbm_res_train = tibble(pred = as.numeric(gbmA.fit.pred), true = as.numeric(contraceptive_pred$contraceptive_freq[trRow]))
gbm_training_err = sqrt(mean((gbm_res_train$true - gbm_res_train$pred)^2))

bart.fit.pred = predict(bart.fit.cv,new_data = x.bart)
bart_res_train = tibble(pred = as.numeric(bart.fit.pred), true = as.numeric(contraceptive_pred$contraceptive_freq[trRow]))
bart_training_err = sqrt(mean((bart_res_train$true - bart_res_train$pred)^2))


############### testing auc

#lr.fit.test.pred = predict(lr.fit,newdata = contraceptive_pred[-trRow,-c(21)])
#sum(lr.fit.test.pred == contraceptive_pred$contraceptive_freq[-trRow])/length(lr.fit.test.pred)

#lasso.fit.test.pred = predict(lasso.fit,newdata = x.lasso.test)
#sum(lasso.fit.test.pred == contraceptive_pred$contraceptive_freq[-trRow])/length(lasso.fit.test.pred)

rf.fit.test.pred = predict(rf.fit,newdata = contraceptive_pred[-trRow,-c(21)])
rf_res_test = tibble(pred = as.numeric(rf.fit.test.pred), true = as.numeric(contraceptive_pred$contraceptive_freq[-trRow]))
rf_test_err = sqrt(mean((rf_res_test$true - rf_res_test$pred)^2))


gbmA.fit.test.pred = predict(gbmA.fit,newdata = contraceptive_pred[-trRow,-c(21)])
gbm_res_test = tibble(pred = as.numeric(gbmA.fit.test.pred), true = as.numeric(contraceptive_pred$contraceptive_freq[-trRow]))
gbm_test_err = sqrt(mean((gbm_res_test$true - gbm_res_test$pred)^2))

bart.fit.test.pred = predict(bart.fit.cv,new_data = x.bart.test)
bart_res_test = tibble(pred = as.numeric(bart.fit.test.pred, digits = 1), true = as.numeric(contraceptive_pred$contraceptive_freq[-trRow]))
bart_test_err = sqrt(mean((bart_res_test$true - bart_res_test$pred)^2))
```

# get bart ranking 

```{r}
props = get_var_props_over_chain(bart.fit.cv) %>% as.data.frame() %>% arrange(desc(.)) %>% rename("prop" = ".") %>% mutate(names = rownames(.))

investigate_var_importance(bart.fit.cv, num_replicates_for_avg = 3)

plot_convergence_diagnostics(bart.fit.cv)


#summary(lm(bart.fit.test.pred ~ true, data = res_1))

rf_coefs = coef(lm(pred ~ true, data = rf_res_test))
gbm_coefs = coef(lm(pred ~ true, data = gbm_res_test))
bart_coefs = coef(lm(pred ~ true, data = bart_res_test))

summary(lm(pred ~ true, data = rf_res_test))
summary(lm(pred ~ true, data = gbm_res_test))
summary(lm(pred ~ true, data = bart_res_test))

rf_res_test%>% ggplot(aes(x = true, y = pred, fill = factor(true))) +
  geom_abline(intercept = rf_coefs[1], slope = rf_coefs[2], color = "blue") +
  geom_boxplot(alpha = .6) + theme_bw() +
  labs(title = "predicted vs actual value") +
  ggsci::scale_fill_lancet()

gbm_res_test%>% ggplot(aes(x = true, y = pred, fill = factor(true))) +
  geom_abline(intercept = gbm_coefs[1], slope = gbm_coefs[2], color = "blue") +
  geom_boxplot(alpha = .6) + theme_bw() +
  labs(title = "predicted vs actual value") +
  ggsci::scale_fill_lancet()

bart_res_test%>% ggplot(aes(x = true, y = pred, fill = factor(true))) +
  geom_abline(intercept = bart_coefs[1], slope = bart_coefs[2], color = "blue") +
  geom_boxplot(alpha = .6) + theme_bw() +
  labs(title = "predicted vs actual value") +
  ggsci::scale_fill_lancet()

mean(rf_res_test$pred)
mean(gbm_res_test$pred)
mean(bart_res_test$pred)
#
#t.test(res[res$true == 1,][,1], res[res$true == 2,][,1])
#t.test(res[res$true == 2,][,1], res[res$true == 3,][,1])


```

# Same models but 10 variables

```{r}
contraceptive_pred_new = contraceptive_pred %>%
  dplyr::select(pdeter, score, depression.x, attitude, enrllyr, marriage, accessibility, single, freeacc, degree, smoking, relationship, drink, contraceptive_freq)

contraceptive_pred_new.male = contraceptive_pred %>%
  filter(sex == "male") %>%
  dplyr::select(pdeter, score, depression.x, attitude, enrllyr, marriage, accessibility, single, freeacc, degree, smoking, relationship, drink, contraceptive_freq)

contraceptive_pred_new.female = contraceptive_pred %>%
  filter(sex == "female") %>%
  dplyr::select(pdeter, score, depression.x, attitude, enrllyr, marriage, accessibility, single, freeacc, degree, smoking, relationship, drink, contraceptive_freq)


x = model.matrix(contraceptive_freq ~ . , contraceptive_pred_new)[,-1] %>% as.tibble() %>%
  as.matrix() %>% .[trRow, ]
x.lasso.test = model.matrix(contraceptive_freq ~ . , contraceptive_pred_new)[,-1] %>% as.tibble() %>%
  as.matrix() %>% .[-trRow, ]
y = contraceptive_pred_new$contraceptive_freq[trRow]

#check dim
dim(x)
length(y)

ctrl = trainControl(method = "cv", number = 5,
                    classProbs = F,
                    #summaryFunction = twoClassSummary,
                    #sampling = "up"
                    )

#lr
#set.seed(123)
#lr.fit_new = train(x = contraceptive_pred_new[trRow,-16], # rm std
#                y = contraceptive_pred_new$contraceptive_freq[trRow],
#                method = "glm",
#                #metric = "ROC",
#                trControl = ctrl
#                ) 
#
##lasso
#lasso.fit_new = train(x,y,
#                method = "glmnet",
#                #metric = "ROC",
#                tuneGrid = expand.grid(alpha = 1,lambda = exp(seq(-7, -6, length #= 100))),
#                trControl = ctrl)
#
#ggplot(lasso.fit_new, highlight = T) +
#  labs(title = "LASSO tuning procedure") +
#  theme_bw()

# rf
set.seed(123)
rf.grid_new = expand.grid(mtry = 2:7,
                      splitrule = "variance",
                      min.node.size = c(100,150,200)
                      )
rf.fit_new = train(contraceptive_freq ~ ., contraceptive_pred_new[trRow,], 
                method = "ranger",
                tuneGrid = rf.grid,
                #metric = "ROC",
                trControl = ctrl)
ggplot(rf.fit_new, highlight = T)  +
  labs(title = "RF tuning procedure") +
  theme_bw()

# Adaboost 
gbmA.grid_new = expand.grid(n.trees = c(100,150,200),
                        interaction.depth = 2:5,
                        shrinkage = c(0.1),
                        n.minobsinnode = c(50,75))


set.seed(123)
gbmA.fit_new = train(contraceptive_freq ~ ., contraceptive_pred_new[trRow,],
                 tuneGrid = gbmA.grid,
                 trControl = ctrl,
                 method = "gbm",
                 distribution = "gaussian",
                 #metric = "ROC",
                 verbose = T)
gbmA.fit$bestTune 
ggplot(gbmA.fit_new, highlight = T) +
    labs(title = "GBM tuning procedure") +
  theme_bw()


set.seed(123)
rf.final_new = ranger(contraceptive_freq ~ ., 
                  contraceptive_pred_new[trRow,], 
                  mtry = rf.fit_new$bestTune$mtry, 
                  splitrule = "variance",
                  min.node.size = rf.fit_new$bestTune$min.node.size,
                  importance = "impurity")

# var importance
vip::vip(rf.final_new,num_features = 23) + theme_bw() + ggtitle("Variable Importance - RF")
vip::vip(gbmA.fit_new$finalModel,num_features = 23) + theme_bw() + ggtitle("Variable Importance - GBM")


#bart
x.bart_new = contraceptive_pred_new[trRow,-14] %>% as.data.frame()
x.bart.test_new = contraceptive_pred_new[-trRow,-14] %>% as.data.frame()
set.seed(123)
y.bart_new = as.numeric(contraceptive_pred_new$contraceptive_freq[trRow])
bart.fit.cv_new = bartMachineCV(X = x.bart_new, y = y.bart_new, k_cvs = 5, verbose = T, num_tree_cvs = c(50,100))
```

#RMSE - 8 variables

```{r}
############### training RMSE
rf.fit.pred_new = predict(rf.fit_new,newdata = contraceptive_pred_new[trRow,-c(14)])
sum(rf.fit.pred_new == contraceptive_pred$contraceptive_freq[trRow])/length(rf.fit.pred)
rf_res_train_new = tibble(pred = as.numeric(rf.fit.pred_new), true = as.numeric(contraceptive_pred$contraceptive_freq[trRow]))
rf_training_err_new = sqrt(mean((rf_res_train_new$true - rf_res_train_new$pred)^2))

gbmA.fit.pred_new = predict(gbmA.fit_new,newdata = contraceptive_pred_new[trRow,-c(14)])
gbm_res_train_new = tibble(pred = as.numeric(gbmA.fit.pred_new), true = as.numeric(contraceptive_pred$contraceptive_freq[trRow]))
gbm_training_err_new = sqrt(mean((gbm_res_train_new$true - gbm_res_train_new$pred)^2))

bart.fit.pred_new = predict(bart.fit.cv_new,new_data = x.bart_new)
bart_res_train_new = tibble(pred = as.numeric(bart.fit.pred_new), true = as.numeric(contraceptive_pred$contraceptive_freq[trRow]))
bart_training_err_new = sqrt(mean((bart_res_train_new$true - bart_res_train_new$pred)^2))


############### testing aRMSE

rf.fit.test.pred_new = predict(rf.fit_new,newdata = contraceptive_pred_new[-trRow,-c(14)])
rf_res_test_new = tibble(pred = as.numeric(rf.fit.test.pred_new), true = as.numeric(contraceptive_pred_new$contraceptive_freq[-trRow]))
rf_test_err_new = sqrt(mean((rf_res_test_new$true - rf_res_test_new$pred)^2))


gbmA.fit.test.pred_new = predict(gbmA.fit_new,newdata = contraceptive_pred_new[-trRow,-c(14)])
gbm_res_test_new = tibble(pred = as.numeric(gbmA.fit.test.pred_new), true = as.numeric(contraceptive_pred_new$contraceptive_freq[-trRow]))
gbm_test_err_new = sqrt(mean((gbm_res_test_new$true - gbm_res_test_new$pred)^2))

bart.fit.test.pred_new = predict(bart.fit.cv_new,new_data = x.bart.test_new)
bart_res_test_new = tibble(pred = as.numeric(bart.fit.test.pred_new, digits = 1), true = as.numeric(contraceptive_pred_new$contraceptive_freq[-trRow]))
bart_test_err_new = sqrt(mean((bart_res_test_new$true - bart_res_test_new$pred)^2))

# summary
## full var training err
training_err = c(rf_training_err, gbm_training_err, bart_training_err)
test_err = c(rf_test_err, gbm_test_err, bart_test_err)
res = rbind(training_err,test_err)
colnames(res) = c("rf", "gbm", "bart")



## 8 var training err
training_err_new = c(rf_training_err_new, gbm_training_err_new, bart_training_err_new)
test_err_new = c(rf_test_err_new, gbm_test_err_new, bart_test_err_new)
res_new = rbind(training_err_new,test_err_new)
colnames(res_new) = c("rf", "gbm", "bart")


res
res_new 


```


# check - 8 variables

```{r}
rf_coefs_new = coef(lm(pred ~ true, data = rf_res_test_new))
gbm_coefs_new = coef(lm(pred ~ true, data = gbm_res_test_new))
bart_coefs_new = coef(lm(pred ~ true, data = bart_res_test_new))
summary(lm(pred ~ true, data = rf_res_test_new))
summary(lm(pred ~ true, data = gbm_res_test_new))
summary(lm(pred ~ true, data = bart_res_test_new))

rf_res_test_new%>% ggplot(aes(x = true, y = pred, fill = factor(true))) +
  geom_abline(intercept = rf_coefs_new[1], slope = rf_coefs_new[2], color = "blue") +
  geom_boxplot(alpha = .6) + theme_bw() +
  labs(title = "predicted vs actual value") +
  ggsci::scale_fill_lancet()

gbm_res_test_new%>% ggplot(aes(x = true, y = pred, fill = factor(true))) +
  geom_abline(intercept = gbm_coefs_new[1], slope = gbm_coefs_new[2], color = "blue") +
  geom_boxplot(alpha = .6) + theme_bw() +
  labs(title = "predicted vs actual value") +
  ggsci::scale_fill_lancet()

bart_res_test_new%>% ggplot(aes(x = true, y = pred, fill = factor(true))) +
  geom_abline(intercept = bart_coefs_new[1], slope = bart_coefs_new[2], color = "blue") +
  geom_boxplot(alpha = .6) + theme_bw() +
  labs(title = "predicted vs actual value") +
  ggsci::scale_fill_lancet()
```


# ordinal logistic

```{r}
library(MASS)
library(generalhoslem)

set.seed(123)
contraceptive_pred_new = contraceptive_pred_new %>% mutate(contraceptive_freq = factor(contraceptive_freq))
contraceptive.polr=polr(contraceptive_freq~ .,data = contraceptive_pred_new)
summary(contraceptive.polr)
lipsitz.test(contraceptive.polr)

# stratified by sex
## male
contraceptive_pred_new.male = contraceptive_pred_new.male %>% mutate(contraceptive_freq = factor(contraceptive_freq))
contraceptive.polr.male=polr(contraceptive_freq~ .,data = contraceptive_pred_new.male)
summary(contraceptive.polr.male)
#lipsitz.test(contraceptive.polr,female)
 
## female
contraceptive_pred_new.female = contraceptive_pred_new.female %>% mutate(contraceptive_freq = factor(contraceptive_freq))
contraceptive.polr.female=polr(contraceptive_freq~ .,data = contraceptive_pred_new.female)
summary(contraceptive.polr.female)
#lipsitz.test(contraceptive.polr,female)

# model prediction
polr.predict.response = predict(contraceptive.polr,contraceptive_pred_new,type= 'class')


# get final table results
## male
plor.table.male = summary(contraceptive.polr.male)$coefficients
p = pnorm(abs(plor.table.male[, "t value"]), lower.tail = FALSE) * 2
OR_point.est = exp(plor.table.male[,1])
OR_ci = exp(c(plor.table.male[,1] - 1.96 * plor.table.male[,2], plor.table.male[,1] + 1.96 * plor.table.male[,2]))
OR_ci_low = OR_ci[1:24]
OR_ci_up = OR_ci[25:48]

plor.table.male = cbind(plor.table.male, "p value" = p)
plor.table.male = cbind(plor.table.male, "point est" = OR_point.est)
plor.table.male = cbind(plor.table.male, "CI_low" = OR_ci_low)
plor.table.male = cbind(plor.table.male, "CI_up" = OR_ci_up)


# intercept
intercept = plor.table.male[-c(1:(nrow(plor.table.male)-2)),]
# final or
write.csv(plor.table.male, "./proportional_odds_male.csv")



## female

plor.table.female = summary(contraceptive.polr.female)$coefficients
p = pnorm(abs(plor.table.female[, "t value"]), lower.tail = FALSE) * 2
OR_point.est = exp(plor.table.female[,1])
OR_ci = exp(c(plor.table.female[,1] - 1.96 * plor.table.female[,2], plor.table.female[,1] + 1.96 * plor.table.female[,2]))
OR_ci_low = OR_ci[1:24]
OR_ci_up = OR_ci[25:48]

plor.table.female = cbind(plor.table.female, "p value" = p)
plor.table.female = cbind(plor.table.female, "point est" = OR_point.est)
plor.table.female = cbind(plor.table.female, "CI_low" = OR_ci_low)
plor.table.female = cbind(plor.table.female, "CI_up" = OR_ci_up)

# intercept
intercept = plor.table.female[-c(1:(nrow(plor.table.female)-2)),]
# final or
write.csv(plor.table.female, "./proportional_odds_female.csv")
```

